### 2024-02-27 index.qmd
write more scaffolding & about bilateral formalism

---
title: "Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer"
authors:
  - name: Roman Ludwig
    orcid: 0000-0001-9434-328X
    email: roman.ludwig@usz.ch
  - name: Yoel Perez Haas
    email: yoel.perezhaas@usz.ch
  - name: Jan Unkelbach
    orcid: 0000-0002-4275-990X
    email: jan.unkelbach@usz.ch
format:
  html:
    code-fold: true
jupyter: python3
execute: 
  cache: true
crossref:
  fig-prefix: "figure"
  tbl-prefix: "table"
  eq-prefix: "equation"
  sec-prefix: "section"
---

# Introduction

- head and neck cancer spreads lymphatically
- may sometimes spread to contralateral side
- spreads more frequently and to larger extent contralaterally when tumor extends the mid-sagittal line
- we describe a model based on previously published hidden Markov model
- we extend it to cover the contralateral side, too
- ipsi- and contralateral side are correlated via time of diagnosis, which is correlated with T-category
- tumor extension over mid-sagittal line is modelled as random variable
- spread probabilities from tumor to contralateral LNLs in case of midline extesnion are linear combinations of these probabilities in case of ipsilateral spread and contralateral spread when tumor is clearly lateralized

# Methods

## Modelling Unilateral Lymphatic Spread

- what we previously did

$$
P \left( \mathbf{X} \mid \mathbf{Z} \right) = \frac{P \left( \mathbf{Z} \mid \mathbf{X} \right) P \left( \mathbf{X} \right)}{P \left( \mathbf{Z} \right)}
$$ {#eq-uni-bayes-law}

Let us for later use define the matrices $\boldsymbol{\Lambda}$:

$$
\boldsymbol{\Lambda} = P \left( \mathbf{X} \mid \mathbf{t} \right) = \begin{pmatrix}
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^0 \\
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^1 \\
\vdots \\
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^{t_\text{max}} \\
\end{pmatrix}
$$ {#eq-lambda-matrix}

The $n$-th row in this matrix corresponds to the distribution over hidden states after $t=n-1$ time-steps.

## Extending the Formalism to Both Sides of the Neck

In analogy to @eq-uni-bayes-law, we can write down the posterior joint probability of an involvement $\mathbf{X}^\text{i}$ ipsilaterally _and_ an involvement $\mathbf{X}^\text{c}$ contralaterally, given a diagnosis of the ipsilateral LNLs $\mathbf{Z}^\text{i}$ and of the contralateral ones $\mathbf{Z}^\text{c}$:

$$
P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \right) = \frac{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right)}{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \right)}
$$

The likelihood of the diagnoses given a hidden state simply factorise: $P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) = P \left( \mathbf{Z}^\text{i} \mid \mathbf{X}^\text{i} \right) \cdot P \left( \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{c} \right)$. And the two factors are contained in the observation matrices $\mathbf{B}^\text{i}$ and $\mathbf{B}^\text{c}$.

The term representing the model's prior probability of hidden involvement does not factorize. However, if we assume that lymphatic spread typically does not cross the mid-sagittal line, we can write it as a factorising sum:

$$
\begin{aligned}
P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) &= \sum_{t=0}^{t_\text{max}} p(t) \cdot P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid t \right) \\
&= \sum_{t=0}^{t_\text{max}} p(t) \cdot P \left( \mathbf{X}^\text{i} \mid t \right) \cdot P \left( \mathbf{X}^\text{c} \mid t \right)
\end{aligned}
$$

This assumption makes intuitive sense: The two sides of the lymphatic network in a typical patient are approximately mirror images of each other. Thus, no major vessels cross the mid-sagittal line. There may, however, be diffusion of lymph fluid accross this line or bulky involvement that redirects lymphatic drainage significantly.

Using this assumption along with @eq-lambda-matrix, we can write the above distribution algebraically as a product:

$$
P \left( \mathbf{X}^\text{i} = \boldsymbol{\xi}_n, \mathbf{X}^\text{c} = \boldsymbol{\xi}_m \right) = \left[ \boldsymbol{\Lambda}^\intercal_\text{i} \cdot \operatorname{diag} p(\mathbf{t}) \cdot \boldsymbol{\Lambda}^\intercal_\text{c} \right]_{n,m}
$$

## Parameter Symmetries

- spread from LNL to LNL probably symmetric

## Tumors Extending Over the Mid-Sagittal Line

- higher prevalence of contralateral involvement when tumor crosses the mid-sagittal plane
- but still lower than on ipsilateral side
- spread to contralateral side in case of midline extension may be describes as linear combination of ipsi- and contralateral spread in case of no midline extension.
- only one additional "mixing parameter" $\alpha$
- tumor's midline extension modelled as a binary RV (lateralized or crossing/extending)

# Results

```{python}
from emcee import backends
from corner import corner
import matplotlib.pyplot as plt
from tueplots import figsizes, fontsizes

backend = backends.HDFBackend(
  filename="models/midline_fix.hdf5",
  read_only=True,
)
samples = backend.get_chain(thin=20)
ndim = samples.shape[-1]

plt.rcParams.update(**figsizes.icml2022_full(nrows=ndim))

fig, axes = plt.subplots(nrows=ndim, sharex=True)
for i, ax in enumerate(axes):
  ax.plot(samples[:,::10,i], color="gray", alpha=0.2)
  ax.axhline(samples[200:,:,i].mean(), color="red")
  ax.set_xlim(0., samples.shape[0])
  ax.set_ylim(0., 1.)
```