### 2025-05-21 manuscript.qmd
extended with first results

---
title: "A mixture of hidden Markov models to predict the lymphatic spread in head and neck cancer depending on primary tumor location"
abstract: |
  Purpose: to be done

  Methods: to be done

  Results: to be done

  Conclusions: to be done
---

# Introduction

Head and neck squamous cell carcinomas (HNSCCs) are known to spread through the lymphatic system often leading to metastases in the lymph nodes [@mukherji_cervical_2001;@shah_patterns_1990]. To minimize nodal recurrences, lymph node levels (LNLs) at risk of harboring occult metastases are typically irradiated electively. Current guidelines for different tumor locations are based on the overall prevalence of nodal disease as reported in literature [@biau_selection_2019;@mukherji_cervical_2001;@shah_patterns_1990].

To personalize the prediction of the risk of occult metastases, given a patient's individual diagnosis, we previously published a large, multi-centric dataset where the lymphatic involvement per LNL is available for each patient[@ludwig_dataset_2022;@ludwig_multi-centric_2023]. Building on this dataset, we introduced an interpretable hidden Markov model (HMM), trained to predict the risk for occult nodal disease, given an individual patient's diagnosis [@ludwig_hidden_2021].

Personalized risk predictions could enable clinicians to safely reduce the elective clinical target volume (CTV-N), potentially decreasing treatment-related side effects that impair a patient's quality of life, without compromising the efficacy of the treatment [@batth_practical_2014].

Initially, separate models were trained for distinct tumor locations, such as the oropharynx and oral cavity. These tumor locations are also used in guidelines to define the elective target volumes [@biau_selection_2019]. However, this approach did not account for variations in lymphatic spread between subsites within these tumor regions. With data from more than 2700 patients available, we can now further analyze subsite specific spread patterns. Closer analysis showed that pooling subsites into a single model led to inaccurate predictions, as it failed to capture distinct lymphatic spread patterns. To resolve this, we propose using a mixture of HMMs, which allows us to model the lymphatic spread more accurately for tumors located near anatomical borders, such as those between the oropharynx and oral cavity (e.g., tumors in the palate).

Additionally, we extend the analysis to a broader mixture model that encompasses tumors of the oral cavity, oropharynx, hypopharynx, and larynx, resulting in further personalized predictions of lymphatic spread across these regions.

# Data on Lymphatic Progression Patterns {#sec-data}

For the analyses in this work, we used seven datasets from 5 institutions resulting in {{< var patients.total >}} patients in total.

1. {{< var patients.USZ_oropharynx >}} oropharyngeal patients from the University of Zurich in Switzerland
2. {{< var patients.CLB_oropharynx >}} oropharyngeal patients from the Centre Léon Bérard in France
3. {{< var patients.ISB_multiside >}} oropharyngeal, larynx and oral cavity patients from the Inselspital Bern in Switzerland
4. {{< var patients.CLB_multiside >}} oropharyngeal, larynx and oral cavity patients from the Centre Léon Bérard in France
5. {{< var patients.USZ_multiside >}} oropharyngeal, larynx and oral cavity patients from the University of Zurich in Switzerland
6. {{< var patients.HVH >}} oropharyngeal patients from the Hospital Vall d'Hebron in Spain (not yet public)
7. {{< var patients.HMC >}} hypopharynx, larynx and oral cavity patients from University Medical Center Groningen (not yet public)

Patients with glottic laryngeal cancer (C32.0) at tumor stages T0 and T1 have been excluded from analysis because these stages, by definition, do not exhibit lymphatic involvement [need source from Panos]. The datasets 1-4 are publicly available as CSV tables [@ludwig_multi-centric_2023;@ludwig_detailed_2022] and can be interactively explored on [LyProX](https://lyprox.org). For each patient the primary tumor subsite is reportd and each indicidual LNL is reported as either metastatic or healthy given the available diagnostic modalities, which include pathology after neck dissection in some patients.
In this work we will stratify the tumor locations into different ICD codes which are depicted in @fig-subsites.

![Anatomical sketch of the tumor subsites and their corresponding ICD-10 codes. Subsite C06 "other parts of mouth" has not been included. Further the The tumor locations are color coded in the following pattern: blue-oral cavity, green-oropharynx, red-hypopharynx, orange-larynx.](figures/Subsites.png){#fig-subsites}

The prevelance of involvement in LNLs I, II, III, IV and V is shown in @fig-involvement. The involvement is stratified per tumor subsite and t-stage. The figure illustrates the variations in LNL involvement between subsites within oral cavity (blue), oropharynx (green), hypopharynx (red) and larynx (orange). The involvement pattern presents a continuous change over the tumor subsites. Where tumors in the oral cavity show the most prominent LNL I involvement. As the tumor location moves towards the oropharynx LNL II involvement increases. Moving the tumor location further in caudal direction towards the hypopharynx increases LNL III involvement while LNL I and II involvement decrease. Larnygeal tumors show the least LNL I involvement.

![Prevalence of ipsilateral LNL involvement stratified by subsite. The subsites are sorted in natural order to represent the continuously changing LNL involvement. The different tumo locations are color coded, where oral cavity subsistes are depicted in blue, larynx in green, hypopharynx in red and larynx in orange. The patient data is further stratified in early t-stage (0-2) and late t-stage (3-4). The legend furter specifies the number of patients in each subsite. For glottis (C32.0) early t-stage only includes T2.](figures/involvement_I_to_V_all_sites_t_staging.png){#fig-involvement}

# Unilateral Model for Lymphatic Progression {#sec-unilateral}

In this chapter we will briefly summarize unilateral model for ipsilateral lymph node involvement introduced in @ludwig_hidden_2021, presenting the notation which is then needed to extend the HMM to a mixture model encompassing multiple tumor subsites.

The HMM describes each LNL $v \in 1, 2, ..., V$ by a binary random variable corresponding to the status of the LNL; healthy (0) or involved (1). The entire state of a patient with $V$ LNLs is defined by the $V$-dimensional vector $\mathbf{X} = [X_1, X_2, ...,X_V]$. In the HMM, a patient's involvement is modeled over time $t$. Thus, a patient's state of lymph node involvement $\mathbf{X}[t]$ evolves over discrete time steps $t$. Let us enumerate all $2^V$ possible states, representing all combinations of LNL involvement. In this paper, we consider ipsilateral LNLs I, II, III, IV and V, which amounts to 32 possible states. The HMM is then specified by a transition matrix $\mathbf{A}$:

$$
\mathbf{A} = \begin{pmatrix} A_{ij} \end{pmatrix} = P \left( \mathbf{X}[t+1] = \boldsymbol{\xi}_j \mid \mathbf{X}[t] = \boldsymbol{\xi}_i \right)
$$ {#eq-transition-matrix}

whose elements $A_{ij}$ contain the conditional probabilities that a state $\mathbf{X}[t]=\boldsymbol{\xi}_i$ transitions to $\mathbf{X}[t+1]=\boldsymbol{\xi}_j$ over one time step. The transition matrix is specified and parameterised via the graphical model shown in @fig-graph. The red arcs in the graph of @fig-graph are associated with the probability that the primary tumor spreads directly to a LNL (parameters $b_v$). The blue arcs describe the spread from an upstream LNL -- given it is already metastatic -- to a downstream level (parameters $t_{v \rightarrow v+1}$).

![Parametrized graphical model of the lymphatic network considering four LNLs. Blue nodes represent the hidden states of LNLs $X_v$, while the red one is the tumor. Arcs represent possible routes of metastatic spread, associated with a probability.](figures/Model_5_LNL.png){#fig-graph width=20%}

Now, let $\boldsymbol{\pi}$ be the _starting distribution_

$$
\boldsymbol{\pi} = \begin{pmatrix} \pi_i \end{pmatrix} = P \left( \mathbf{X}[0] = \boldsymbol{\xi}_i \right)
$$ {#eq-starting-distribution}

denoting the probability to start in state $\boldsymbol{\xi}_i$ at time step 0. Assuming that every patient started with all LNLs being healthy, we set $\pi_i$ to zero for all states  except the completly healthy state $\boldsymbol{\xi} = \begin{pmatrix} 0, 0, 0, 0, 0 \end{pmatrix}$, which has probability one.

Using the quantities introduced so far, the probability $P \left( \mathbf{X}[t]=\boldsymbol{\xi}_i \right)$ to be in state $\boldsymbol{\xi}_i$ in time step $t$ can now be conveniently expressed as a matrix product:

$$
P \left( \mathbf{X}[t]=\boldsymbol{\xi}_i \right) = \left( \boldsymbol{\pi} \cdot \mathbf{A}^t \right)_i
$$ {#eq-evolution}

This evolution implicitly marginalizes over all possible paths to arrive at state $\boldsymbol{\xi}_i$ after $t$ time-steps. Additionally, we must marginalize over the unknown time of diagnosis using a time-prior $P_T(t)$ which is defined by a binomial distribution. The t-stage of the tumor can be included in the model by choosing different parameterizations of the binomial distribution, considering that a tumor in late t-stages was diagnosed later than a tumor in early t-stages, therefore shifting the probability of diagnosis to later time steps. This finally defines the probability distribution over all states of lymph node involvement.

$$
P \left( \mathbf{X}=\boldsymbol{\xi}_i \mid \boldsymbol{\theta}, \mathbf{T} \right) = \sum_{t=0}^{t_\text{max}} P_T(t) \left( \boldsymbol{\pi} \cdot \mathbf{A}^t \right)_i
$$ {#eq-marginalized-evolution}

where $\boldsymbol{\theta}=\{ b_v, t_{v \rightarrow v+1} \}$ denotes the set of all model parameters (7 in our case). Fortunately, the exact length and shape of this distribution has little impact as previously shown [@ludwig_hidden_2021]. We set $t_\text{max}=$ {{< var model.max_t >}} and $P_\text{early}(t)$ to a binomial distribution with parameter {{< var model.first_binom_prob >}}. Further details on the HMM can be found in @ludwig_hidden_2021 and @zora231470.

With @eq-marginalized-evolution we can compute the probability of a patient being in any state $\xi_i$. 
To train the model, we assume that the observed diagnoses in our dataset $\mathbf{D}$ directly reflect the underlying hidden states $\mathbf{X}$ of the patient. Thus, learning the model parameters corresponds to maximizing the probability of observing the dataset $\mathbf{D}$:

$$
P \left( \mathbf{D} \mid \boldsymbol{\theta} \right) = \prod_k^K P \left( \mathbf{X}_k=\boldsymbol{\xi}_i \mid \boldsymbol{\theta}, T_k \right)
$$ {#eq-likelihood-HMM}

In @eq-likelihood-HMM we compute the likelihood of observing the diagnosis of each patient $k$, i.e. t-stage $T_k$ and involvement  $\mathbf{X}_k=\boldsymbol{\xi}_i$.

## Risk Estimation
The goal of the HMM is to predict the risk of occult metastases in LNLs, given a patient's diagnosis. In this framework this corresponds to computing the probability of a hidden state $\mathbf{X}=\boldsymbol{\xi}_i$ given the observed diagnosis $\mathbf{Z}$ and the model parameters $\boldsymbol{\theta}$. With Baye's theorem we can compute this probability as follows:

$$
P(\mathbf{X} =\boldsymbol{\xi}_i \mid \mathbf{Z}, \boldsymbol{\theta}) = \frac{P(\mathbf{Z} \mid \mathbf{X}=\boldsymbol{\xi}_i, \boldsymbol{\theta}) P(\mathbf{X}_k=\boldsymbol{\xi}_i \mid \boldsymbol{\theta})}{P(\mathbf{Z} \mid \boldsymbol{\theta})}
$$ {#eq-bayes}

where $P(\mathbf{Z} \mid \boldsymbol{\theta})$ is the marginal likelihood of the data, which can be computed by summing over all possible hidden states $\boldsymbol{\xi}_i$:

$$
P(\mathbf{Z} \mid \boldsymbol{\theta}) = \sum_{\boldsymbol{\xi}_i} P(\mathbf{Z} \mid \boldsymbol{\xi}_i, \boldsymbol{\theta}) P(\boldsymbol{\xi}_i \mid \boldsymbol{\theta})
$$ {#eq-marginal-likelihood}

With this we can compute the probability for each hidden state and subsequently we can compute the probability for each LNL $v$ to be involved given the diagnosis $\mathbf{Z}$ and the model parameters $\boldsymbol{\theta}$ by summing over all possible hidden states $\boldsymbol{\xi}_{i,v}$ where the LNL $v$ is involved:

$$
P(X_v=1 \mid \mathbf{Z}, \boldsymbol{\theta}) = \sum_{\boldsymbol{\xi}_{i,v}} P(\mathbf{X} =\boldsymbol{\xi}_i \mid \mathbf{Z}, \boldsymbol{\theta})
$$ {#eq-risk}

As opposed to the model training, we do not assume anymore that the diagnoses correspond to the hidden states. Instead, we consider the sensitivity and specificity of the diagnostic modalities used to determine the involvement of LNLs. Enabling the uncertainty originating from diagnostic modalities to be considered in the model. Here we chose the sensitivity of 0.81 for clinical modalities corresponding to the literature value for imaging modalities [@de_bondt_detection_2007] and specificity of 1.0, indicating that all positively diagnosed LNLs are indeed involved.

# Mixture Model for Lymphatic Spread {#sec-mixture}

Primary tumors at different subsites exhibit distinct lymphatic spread patterns. This presents a challenge when attempting to generalize predictive models across subsites. One approach, as introduced in [@ludwig_dynamic_2021], uses a Hidden Markov Model (HMM) trained specifically for oropharyngeal cancer. However, extending this model to other subsites would either require generalizing over several subsites or training a separate model for each. The former approach sacrifices precision, particularly for subsites with fewer patients, while the latter approach becomes computationally intensive and introduces large uncertainties for subsites with limited patient data, such as C04 (Floor of mouth) or C05 (Palate).

To address these challenges and exploit the anatomical similarities between nearby subsites, we introduce a mixture model that combines data from all subsites into a single model. This model accounts for anatomical proximities, thereby improving predictive power while maintaining computational efficiency.

## Mixture Model Formulation

The mixture model assumes that the data is generated by a set of $M$ different lymphatic spread models. Each patient $k$, with their primary tumor in subsite $s \in (1, 2, \ldots, S)$, is assumed to be generated by one specific model $m \in (1, 2, \ldots, M)$ from this set of $M$ models with probability $\pi_m^s$. These so-called _mixing parameters_ $\boldsymbol{\pi}^s = \{\pi_1^s, \pi_2^s, \dots, \pi_M^s\}$ must satisfy the condition

$$
\sum_{m}^M \pi_m^s = 1, \quad \forall s
$$

If we could record the componet $m$ from which a patient $k$ was drawn from, we could store this information in a binary latent vector $\boldsymbol{\epsilon}_k$. The vector $\boldsymbol{\epsilon}_k$ has length $M$, with exactly the $m$-th element set to 1, indicating which model generated the patient’s data, and all other elements set to 0. Thus, for patient $k$, the latent variable $\boldsymbol{\epsilon}_k$ can be interpreted as a categorical indicator variable that encodes the assignment to one of the $M$ lymphatic spread models. Typically in mixture models, this so-called _latent variable_ is unknown. However, it can be inferred and is useful for inferring the models' parameters.

The joint probability of the observed data $\mathbf{D}$ (i.e., patient data) and the latent variables $\boldsymbol{\epsilon}$ - sometimes called the _complete data likelihood_ - is given by:

$$
P(\mathbf{D}, \boldsymbol{\epsilon} \mid \boldsymbol{\theta}, \boldsymbol{\pi}) = \prod_{k}^{K} \prod_{m}^{M} \left[ \pi_m^{s_k} P(\mathbf{D}_k \mid \boldsymbol{\theta}_m) \right]^{\epsilon_{k}^m}
$$ {#eq-complete-likelihood}

Here:

- $\pi_m^{s_k}$ is the mixing coefficient for subsite $s_k$ (where patient $k$ has their tumor) and model $m$,
- $P(\mathbf{D}_k \mid \boldsymbol{\theta}_m)$ is the likelihood of patient $k$'s diagnosis, i.e. The involvement state $\mathbf{X}_k = \boldsymbol{\xi}_i$ and t-stage $T_k$, given that it was generated by model $m$,
- $\boldsymbol{\theta}_m$ represents the parameters of model $m$,
- $\epsilon_{k}^m$ is the latent variable that indicates patient $k$ was generated by model $m$.

In contrast, the _incomplete data likelihood_ marginalizes over all possible latent assignments to reflect the uncertainty about which model generated each patient’s data:

$$
P(\mathbf{D}, \mid \boldsymbol{\theta}, \boldsymbol{\pi}) = \prod_{k}^{K} \sum_{m}^{M} \pi_m^{s_k} P(\mathbf{D}_k \mid \boldsymbol{\theta}_m)
$$ {#eq-incomplete-likelihood}

Ultimately, we want to find the parameters which maximize this likelihood function for the given data.

Note the summation inside the product. This structure of mixture models makes naive inference difficult, because the logarithm of this quantity is expensive to compute and not easy to differentiate. Thus, inferring the latent assignment is helpful, because the complete data (log-)likelihood (@eq-complete-likelihood) does not suffer from this shortcoming.

To infer the latent assignment $\boldsymbol{\epsilon}_k$, we start with its distribution, given the observed data and model parameters:

$$
\gamma(\epsilon_k^m) := \mathbb{E}[\epsilon_k^m] = \frac{\pi_m^{s_k} P \left( \mathbf{D}_k \mid \boldsymbol{\theta}_m \right)}{\sum_{j \leq M} \pi_j^{s_k} P \left( \mathbf{D}_k \mid \boldsymbol{\theta}_j \right)}
$$ {#eq-responsibilities}

This expectation value - often called the _responsibility_ - describes the probability that patient $k$ was generated by model $m$. It can be used to compute the expected complete data likelihood:

$$
\mathbb{E}_{\boldsymbol{\epsilon}} \left[ P \left( \mathbf{D}, \boldsymbol{\epsilon} \mid \boldsymbol{\theta}, \boldsymbol{\pi} \right) \right] = \prod_{k=1}^K \prod_{m=1}^M \left[ \pi_m^{s_k} P \left( \mathbf{D}_k \mid \boldsymbol{\theta}_m \right) \right]^{\gamma(\epsilon_k^m)}
$$ {#eq-expected-complete-likelihood}

This expected complete data likelihood has the same tractable form as @eq-complete-likelihood and thus allows us to infer both the mixing coefficients $\boldsymbol{\pi}$ as well as each component model's parameters $\boldsymbol{\theta}_m$ using straightformward inference. Also, these two sets of parameters are the only ones used for the later risk prediction. The responsiblities $\gamma(\epsilon_k^m)$ are only used during inference.

In @fig-model-simple the mixture coefficients are illustrated. Subsites with different spread patterns, such as Gum (C03) and Base of tongue (C01), are expected to get different model assignments. Nonetheless, the latent variables for two patients with the same diagnosis, i.e. same involvement $\boldsymbol{\xi}$ and and t-stage $T$, but different subsites are the same.

![Illustration of mixture parameter assignment. Since Gum and Base of tongue express different spread patterns, the two models are expected to have different model assignments. The arrow visibility represents the value of the mixture parameter $\pi$, where the more visible the arrow, the larger the value for $\pi$](figures/mixture_model_simplified.png){#fig-model-simple}

It is important to note that this mixture model algorithm diverges from the conventional formulation of mixture models as introduced in @bishop_pattern_2006. In Bishop's work, the mixture model is defined with a single set of mixture parameters $\boldsymbol{\pi}$, implying that all data points are generated from the same mixture of models. In our approach, however, we categorize the data points (patients) into distinct cohorts (subsites), with each subsite characterized by its own set of mixture parameters. Consequently, we construct $S$ separate mixture models, each possessing different mixture parameters while all models share the same underlying probability distributions, specifically the varying lymphatic spread models.

## Expectation-Maximization (EM) Algorithm

To find the mixing coefficients $\boldsymbol{\pi}$ and all models' parameters $\boldsymbol{\theta}_m$ that maximize @eq-incomplete-likelihood, we follow an iterative approach called _expectation-maximization_ or _EM-algorithm_. With arbitrarily initialized starting parameters, we alternate between the following two steps:

1. In the **E**xpectation step, we compute the responsibilities $\gamma(\epsilon_k^m)$, which represent the probabilities of a patient $k$ originating from one of the models $m$, given the current estimates of $\boldsymbol{\theta}$ and $\boldsymbol{\pi}$, as given in @eq-responsibilities.
2. During the **M**aximization step, we find a new set of parameters that maximize the new expected complete data likelihood (@eq-expected-complete-likelihood). For the mixture coefficients, we can even find an analytic solution to the new maximum:
   $$
   \pi_m^s = \frac{1}{|K_s|} \sum_{k \in K_s} \gamma(\epsilon_k^m)
   $$
   where we sum over the set of all patients $K_s$ with their tumor in subsite $s$.\
   The models' new parameters are found by numerically maximizing the respective likelihood, weighted by the responsibilities:
   $$
   \ln P(\mathbf{D}, \boldsymbol{\epsilon} \mid \boldsymbol{\theta}_m) = \sum_k^K \gamma(\epsilon^m_{k}) \big[ \ln \pi_m^{s_k} + \ln P \left( \mathbf{D}_k \mid \boldsymbol{\theta}_m \right) \big]
   $$ {#eq-maximization-step}

By iterating these steps, the EM algorithm is guaranteed to converge to a (local) maximum of the incomplete data likelihood (@eq-incomplete-likelihood).

## Uncertainty estimation {#sec-uncert}

To estimate the uncertainty associated with both the model parameters and the resulting risk estimates, we employ a bootstrapping approach. Although more comprehensive sampling methods exist for the EM algorithm—such as the imputation-posterior (IP) algorithm, which samples from both model parameters and latent variables—we opt for a computationally more efficient alternative that still yields multiple estimates of all parameters.

Our resampling procedure involves the following steps:

1. Resample the observed data with replacement, maintaining the original number of samples per ICD code.

2. Fit the model to each resampled dataset.

3. Extract the model parameters and risk estimates from each fitted model.

This method enables a straightforward uncertainty analysis while avoiding the computational cost of more intensive sampling techniques.

# Model Training Evaluation {#sec-model_train}
In the following sections we will analyze the model training for three components first, considering Oral Cavity, Oropharynx and Hypopharynx data. Then we extend the model to four components, additionally including Larynx data. Both models will include LNLs I, II, III, IV and V. LNLs VI and VII are omitted for simplicity and since the prevalence for these LNLs never exceeds 11% (@fig-involvement), meaning that the risk of involvement is low in most cases.

## Three Component Mixture Model {#sec-3comp}

We first evualuate the methodology for a mixture model with M = 3 components. We include the ICD codes as subsites for oral cavity, hypopharynx and oropharynx. In @fig-convergence the convergence of the negative log-likelihood and change in model parameters is depicted. After a random inizialization, the algorithm rapidly converges. The algorithm was stopped when the difference of log-likelihood between two iterations was below 0.01.

![The y-axis on the left shows the negative likelihood convergence depicted in the blue line. The y-axis on the right shows the sum of absulte difference between all model parameters showing that the parameter values stabilize rapidly as well.](figures/Convergence_3_comp.png){#fig-convergence}

In @fig-3_simplex, we visualize the resulting mixture coefficients $\boldsymbol{\pi}$ using a spatial representation, where the vertices of the triangle correspond to the three components. In @fig-3-matrix, these mixture coefficients are presented in matrix form, with the y-axis representing the ICD codes, showing how the mixture components in each row add up to 1.

The spatial plot in @fig-3_simplex illustrates how the model assigns the three components to different tumor subsites. Component 0, located at the bottom right of the triangle, primarily characterizes hypopharyngeal subsites. Both hypopharyngeal subsites, piriform sinus (C12) and hypopharynx (C13), are strongly assigned to this component.
Similarly, component 1, located at the bottom left, characterizes oral cavity subsites. For instance, the gum (C03), which has the highest LNL I involvement is fully assigned to this subsite. As subsites anatomically approach the oropharynx, their mixture coefficients for the oropharynx-like component, at the top of the triangle, increase. This is evident in the subsites C02 (tongue) and C05 (palate), which display a higher proportion of oropharyngeal influence in their mixture. These results conform well with the involvement patterns observed in the data in figure @fig-involvement.
The base of tongue subsite (C01), which exhibits the highest involvement of LNL II, is fully assigned to oropharynx-like component. Similarly, subsite C10, which includes several oropharyngeal regions, is assigned roughly 50% to the oropharynx-like component, with the remaining mixture distributed across the other two components, representing its large variation in spread patterns.

::: {layout="[75,30]" layout-valign="bottom"}

![Assignment of each subsite to each of the three components. The closer a subsite is to a vertex, the more it is assigned to the corresponding component, with component 0 on the bottom right, 1 on the bottom left and 2 on the top. The size of the marker (area) corresponds to the number of patients in each subsite.](figures/mixture_components_3_simplex_t_stages_new.png){#fig-3_simplex}

![Matrix representation of component assignment. Each row of the matrix corresponds to each ICD code. The collums represent the three different components](figures/mixture_components_3_t_stages_new.png){#fig-3-matrix}
:::

To evaluate the uncertainty in the model parameters, the bootstrapping approach introduced in @sec-uncert has been applied. Specifically, we generated 200 bootstrap samples by resampling the observed data with replacement, maintaining the original number of samples per ICD code. For each bootstrap sample, the mixture model was re-fitted, and the resulting parameters were recorded. This process yielded a distribution of parameter estimates, allowing us to compute confidence intervals and assess the variability in the model's predictions.

The results of the bootstrapping analysis are shown in @fig-simplex_uncert and @fig-3_comp_scatter_uncert. The figure illustrates the variability in the mixture coefficients $\boldsymbol{\pi}$ for each subsite in the same representation as @fig-3_simplex. The figure shows that the uncertainties match the continuity of the spread patterns. The base of tongue (C01) has most of its uncertainty along in between the oropharynx component and the hypopharynx component. Similarly, the hypopharynx subsite (C13) mostly presents uncertainty between these two components. The more general oropharnyx subsite (C10) shows more uncertainty in each components direction, due to its mixed spread pattern. The subsite for other pars of mouth (C03) however, only has uncertainty along the axis connecting the oropharyngeal component and the oral cavity componen, therefore never mixing with the hypopharynx-like component on the right. In figure @fig-3_comp_scatter_uncert the asymmetry of the uncertainty is better depicted. E.g. the base of tongue (C01) subsite and the hypopharynx (C13) subsite have strongly asymmetric uncertainties. Further, the plot better characterizes the uncertainty in the palate subsite (C05), where the mixture and uncertainty is mainly in the oropharynx-like and the oral cavity-like components if we center the uncertainty around the optimal values instead of the mean values as it is depicted in @fig-simplex_uncert.

![Variability in mixture coefficients $\boldsymbol{\pi}$ across bootstrap samples for each subsite. Where the dots represent the optimal values and the ellipses represent the 68% percentile centered around the mean value. The 68% percentile are computed assuming a 2D-gaussian distribution. However, with the bounds of 0 and 1 the sampled mixture coefficients do not exactly follow a gaussian distribution.](figures/mixture_coefficients_uncert_simplex.png){#fig-simplex_uncert}

![Variability in mixture coefficients $\boldsymbol{\pi}$ across bootstrap samples for each subsite. The dots represent the optimal values. The bars include 68% of all sample values centered around the optimal values.](figures/3_comp_mixture_coefficients_uncert_scatter.png){#fig-3_comp_scatter_uncert}

Additionally, the bootstrapping approach enables us to quantify the uncertainty in the model predictions for each subsite and LNL. To benchmark the model we analyze the predicted and observed prevalences for LNLs I, II, III, IV, and V. These predictions are stratified by tumor subsite and t-stage, providing a comprehensive view of the model's performance and its associated uncertainty.

In @fig-oral_cavity we present the predictions for the oral cavity subsites C03 and C05. Here we can clearly see in figure @fig-oral_cavity_C03_early and @fig-oral_cavity_C03_late that the difference in prevalence is substantial, even though both subsites are grouped in the same tumor location. The model predicts a higher prevalence for LNL II in subsite C05 (palate) compared to subsite C03 (gum). This is consistent with the observed data, where the prevalence of LNL II involvement is higher for the palate than for the gum. However, the model only predicts a slightly higher prevalence for LNL I in subsite C03 compared to subsite C05. This is not consistent with the observed data, where the prevalence of LNL I involvement is significantly higher for the gum than for the palate. Thus, showing a limitation in the model to properly split these different involvements. In late t-stages, depicted in figure @fig-oral_cavity_C03_late and @fig-oral_cavity_C05_late, the model predicts a higher prevalence for LNL II in subsite C05 compared to the early t-stage, but not fully capturing the increase in prevalence. For subsite C03 the LNL II involvement is predicted to increase  as well, capturing the increase in prevalence. However, for LNL I, while capturing the increase in prevalence, it does not optimally predict the prevalence, similarly to the early t-stage. For LNLs III, IV and V, the model consistently predicts the low prevalences observed in the data for both subsites.

::: {#fig-oral_cavity layout-ncol=1}

![](figures/3_predicted_prevalence_C03_early.png){#fig-oral_cavity_C03_early}

![](figures/3_predicted_prevalence_C05_early.png){#fig-oral_cavity_C05_early}

![](figures/3_predicted_prevalence_C03_late.png){#fig-oral_cavity_C03_late}

![](figures/3_predicted_prevalence_C05_late.png){#fig-oral_cavity_C05_late}

Observed prevalence (solid lines) and predicted prevalence (histograms) for subsites C03 and C05. Figures a and b show the prevalences for early t-stages. Figures c and d show the prevalences for late t-stages. 
:::

In @fig-oropharynx we present the predictions for the oropharynx subsites C01 and C10. 
In figures @fig-oropharynx_C01_early and @fig-oropharynx_C10_early we can see that the model predicts a higher prevalence for LNL II in subsite C01 (base of tongue) compared to subsite C10 (oropharynx). This is consistent with the observed data. In figures @fig-oropharynx_C01_late and @fig-oropharynx_C10_late we can see that the model precisely captures the increased prevalence in late t-stage for both subsites. LNLs I, IV and V are predicted to have low prevalences for both subsites, but the model does not perfectly fit the predicted prevalences to the observed prevalences.

::: {#fig-oropharynx layout-ncol=1}

![](figures/3_predicted_prevalence_C01_early.png){#fig-oropharynx_C01_early}

![](figures/3_predicted_prevalence_C10_early.png){#fig-oropharynx_C10_early}

![](figures/3_predicted_prevalence_C01_late.png){#fig-oropharynx_C01_late}

![](figures/3_predicted_prevalence_C10_late.png){#fig-oropharynx_C10_late}

Observed prevalence (solid lines) and predicted prevalence (histograms) for subsites C01 and C10. Figures a and b show the prevalences for early t-stages. Figures c and d show the prevalences for late t-stages. 
:::

In summary, the mixture model successfully distinguishes between subsites, capturing the continuous variation in lymphatic spread patterns across tumor locations. However, the model's predictions are not always perfect. While it generally aligns with observed data, certain nuances, such as the higher prevalence of LNL I involvement in subsite C03 compared to C05, are not fully captured.

## Four Component Mixture Model {#sec-4comp}

We can extend the mixture model to include the larynx. The larynx patients are more finely divided into ICD codes C32.0, C32.1 and C32.2 as there is a notable difference between these ICD codes in @fig-involvement.

Simlarly to the three component model, we can analyze the convergence over the iterations of the EM-algorithm. In @fig-convergence4 we can see that in this more complex model, the likelihood space becomes more complex as at around 200 iterations, the negative log-likelihood starts to increase faster again.

![The y-axis on the left shows the negative likelihood convergence depicted in the blue line. The y-axis on the right shows the sum of absulte difference between all model parameters.](figures/convergence_4_components_4_loc.png){#fig-convergence4}

The component assignment is shown in @fig-4-matrix. Similarly to the 3-component model the different tumor locations are assigned to a one of the components. In this case, the new component is used to model laryngeal subsites.

![Matrix representation of component assignment. Each row of the matrix corresponds to each ICD code. The collums represent the three different components](figures/mixture_components_4_t_staging_add_larynx.png){#fig-4-matrix}

add some analysis (level specific predictions and also comparison to single tumor location models.)

# Model Predictions {#sec-model_pred}

Given the model parameters, we can compute the risk of occult metastases given a patient's diagnosis. The model can be applied to predict the risk of involvement for each LNL, given the patient's t-stage and primary tumor subsite.

The risk predictions are computed by applying the @eq-risk to each component and subsequently mixing the results using the mixture coefficients $\boldsymbol{\pi}$. The risk of involvement for each LNL $v$ in subsite $s$ is given by:
$$
P(X_v^s=1 \mid \mathbf{D}, \boldsymbol{\theta}) = \sum_{m=1}^M \pi_m^{s} P(X_v=1 \mid \mathbf{D}, \boldsymbol{\theta}_m)
$$ {#eq-risk-mixing}

where $s$ is the subsite and $\boldsymbol{\theta}_m$ are the parameters of model component $m$.

For all estimates we will set a 5% threshold for the risk of involvement. This means that if the model predicts a risk of involvement of 5% or higher, we will consider the LNL to be at a "high" risk of involvement and should be irradiated. Consequently all LNLs with a risk of involvement below 5% will be considered to be at a "low" risk of involvement and should not be irradiated. This allows us to compare the model predictions with the current clinical guidelines for elective irradiation of LNLs.

In the following sections we will analyze the model predictions for the four component model only and for selected clinical diagnoses.

## Oral cavity

Oral cavity tumors typically spread to LNLs I, II and III as shown in @fig-involvement. Consequently, the clinical guidelines recommend irradiating LNLs I, II, III for all tumors, irrespective of t-stage and even if there is no LNL involvement [@biau_selection_2019].

We first analyze the risk predictions for the case where we diagnosed